\LARGE \bf{Hall C 12 GeV Trigger Set-Up}

### 2019-07-17 MY_THESIS/HallC_12GeV_Trigger/version2/trigger_v2.tex
updated trgger documentation, added hall c analysis procedures documentation

\usepackage{url}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{graphicx}
\usepackage[scientific-notation=true]{siunitx}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage[top=0.6 in, bottom=0.7 in, left=0.7in, right=0.7in]{geometry}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{vector}
\usepackage{amsmath}
\usepackage{empheq}
\usepackage{amssymb}
\usepackage{tgtermes}
\usepackage{gensymb}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{epstopdf}
\usepackage[singlelinecheck=false]{caption}
\usepackage{datetime}
\usepackage[hyperfootnotes=false, pdfnewwindow, colorlinks=true]{hyperref}

\AtBeginDocument{% optionally set colors to your liking
  \hypersetup{
    urlcolor=red,
    citecolor=green,
    linkcolor=blue,
  }%
}

@@

\newcommand{\shodthrs}{-30 mV }
\newcommand{\quartzthrs}{-60 mV }
\newcommand{\shodgate}{60 ns } 
\newcommand{\sngcthrs}{-50 mV}
\newcommand{\sngcgate}{30 ns}
\newcommand{\shgcthrs}{-50 mV}
\newcommand{\shgcgate}{30 ns}
\newcommand{\saerthrs}{-50 mV}
\newcommand{\saergate}{30 ns}

\title{\Large \bf{Hall C 12 GeV Trigger Set-Up}}

\author{Carlos Yero}

\date{\today}
\maketitle

@@

\subsection{HMS Detector Hut}
The HMS Drift Chambers are read out through a VXS Crate (ROC3) in the detector hut electronics rack. The signals are carried through 16-channel ribbon cables
which are fed in various CAEN1190 TDC modules. The \textbf{TI} (or Trigger Interface) module at the front end of the crate distributes the readout
trigger throughout all modules in the crate and initiates data readout.
\begin{figure}[h!]
  \centering
  \includegraphics[width=5.0in, height=4.0in]{HMS_stack.pdf}
  \caption{HMS detector stack.}
  \label{fig:hms_stack}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=5.0in, height=4.0in]{HMS_Hut_Rack.pdf}
  \caption{HMS detector hut electronic rack and patch panels.}
  \label{fig:hms_hut_rack}
\end{figure}
The rest of the HMS detector signals(Gas \c{C}erenkov, Hodoscope, Calorimeter) are sent to the Hall C Floor Patch Panel via the hut Patch, with the exception
of the Aerogel, which is sent directly from the detector to the Floor Patch.

\newpage
All the signals are then sent to the Counting Room Patch Panel to be processed
by the electronics. (See Figure \ref{fig:hms_patch})

\begin{figure}[h]
  \centering
  \includegraphics[width=6.0in, height=2.0in]{HMS_patch.pdf}
  \caption{HMS Patch Diagram from detectors to Counting Room.}
  \label{fig:hms_patch}
\end{figure}

\subsection{SHMS Detector Hut}
Similarly to the HMS Drift Chambers, the SHMS Drift Chambers are also read out by TDCs in a VXS Crate in the SHMS electronics hut (See Figure \ref{fig:shms_hut_rack2}).
The Shower Counter 224 signals are fed directly into the Flash ADC (fADC-250) modules in a VXS Crate (ROC4). The Pre-Shower signals (x14/side) pass
through a 50:50 splitter and a part is fed into fADCs. The other part is partially summed in the hut and sent via the hut patch panel to the
Counting Room Patch (See Figure \ref{fig:shms_hut_rack1}). 
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.0in, height=4.0in]{SHMS_stack.pdf}
  \caption{SHMS detector stack.}
  \label{fig:shms_stack}
\end{figure}
\begin{figure}[h!]
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=3.0in, height=4.0in]{SHMS_Hut_Rack01.pdf}
  \captionof{figure}{SHMS Hut Patch Panel (left) and Electronics Rack (right).}
  \label{fig:shms_hut_rack1}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=2.0in, height=4.0in]{SHMS_Hut_Rack02.pdf}
  \caption{SHMS Drift Chamber Electronics Rack.}
  \label{fig:shms_hut_rack2}
\end{minipage}
\end{figure}
\newpage
The rest of the SHMS detector signals (HGC/NGC, Hodoscope, Aerogel) are sent to the Counting Room via the hut patch panel (See Figure \ref{fig:shms_patch}). 
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.0in, height=2.0in]{SHMS_patch.pdf}
  \caption{SHMS Patch Diagram from detectors to Counting Room.}
  \label{fig:shms_patch}
\end{figure}

\newpage
\subsection{Hall C Counting Room}
Once the detector signals arrive at the Counting Room Patch (See Figure \ref{fig:cr_patch}), they are processed by the NIM/CAMAC electronics (See Figure \ref{fig:cr_racks}) to form the single arm and coincidence triggers
for each spectrometer. The signals are also sent to ADCs/TDCs to determine energy and timing information for individual detectors as well as trigger TDC information.

\begin{figure}[h!]
  \centering
  \includegraphics[width=3.0in, height=4.0in]{CR_Patch.pdf}
  \caption{Counting Room Patch Panels.}
  \label{fig:cr_patch}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.3\textwidth}
  \centering
  \includegraphics[width=2.0in, height=4.0in]{CR_rack4.pdf}
\end{subfigure}%
\begin{subfigure}{.2\textwidth}
  \centering
  \includegraphics[width=2.0in, height=4.0in]{CR_rack5.pdf}
\end{subfigure}
\caption{Counting Room main Electronic Racks for HMS/SHMS detectors.}
\label{fig:cr_racks}
\end{figure}

\section{HMS Trigger Set-Up} 
\indent The XY scintillator arrays (hodoscope planes) will form part of the standard HMS trigger configuration.  
Additional particle detectors may also be incorporated into the HMS trigger as required by different experiments. The Gas \v{C}erenkov and Calorimeter triggers will be
used for e/$\pi$ separation, whereas the Aerogel \v{C}erenkov trigger will be used for $\pi$/K/p separation.

\subsection{Hodoscope Pre-Trigger}

\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=3.2in]{HMS_HODO_TRIGGER.pdf}
  \caption{HMS Hodoscope Electronics Diagram.}
  \label{fig:hms_hod_trg}
\end{figure}

\indent Each hodoscope plane consists of an array of scintillator bars coupled to a PMT at each end (See Figure \ref{fig:hms_stack}), so each bar reads out two signals. As shown in Figure \ref{fig:hms_hod_trg},
for example, hodoscope plane h1X consists of 32 signals (16 bars) read out in the Counting House (CH) patch. Each side of the plane (x16 signals/side) is fed into a 64-channel input passive splitter (16 Ch./set).
One-third (33\%) of the signal amplitude is sent via a 16-channel ribbon cable to a 64 Ch. input Ribbon-to-BNC converter (16 Ch./set) which outputs are fed into a 16-channel NIM input flash ADC (fADC). The remaining two-thirds (66\%)
of the signal amplitude is sent to a 16-Ch. input CAMAC Discriminator unit. The HMS discriminators thresholds and gate widths were set to \hhodthrs and \hhodgate, respectively.\\
\indent The discriminated signals are sent via two ribbon-cable outputs to CAEN1190 TDCs/Scalers (daisy-chained) and to a LeCroy 4564 CAMAC Logic unit to form the plane pre-triggers.
The Logic Unit takes four sets of 16-Ch. input ribbon cable and forms a 16-fold OR for each set by default. Further boolean operations are done through the module backplane by connecting a twisted pair cable to the pin corresponding to the
desired boolean operation. For hodoscope plane pre-triggers, the boolean operations are as follows:
\begin{empheq}[box=\fbox]{align*}
& \text{h1X = h1X+ (16-fold OR) \textit{AND} h1X- (16-fold OR)} \\ 
& \text{h1Y = h1Y+ (10-fold OR) \textit{AND} h1Y- (10-fold OR)} \\
& \text{h2X = h2X+ (16-fold OR) \textit{AND} h2X- (16-fold OR)} \\ 
& \text{h2Y = h2Y+ (10-fold OR) \textit{AND} h2Y- (10-fold OR)} 
\end{empheq}

\indent Once a pre-trigger has been made for each plane, they are sent to a NIM/ECL converter (Level Translator - Phillips Scientific (or P/S) Model 7126) via twisted pair cables to convert the ECL signal (twisted pair)
to a NIM signal. The NIM output is then sent to individual sets of a P/S Model 752 NIM Logic unit to adjust the widths of each of the plane pre-triggers as necessary before making a coincidence. An X-Y hodoscope plane
coincidence (h1 = h1X \textit{AND} h1Y, h2 = h2X \textit{AND} h2Y) is then made by feeding each hodoscope X-Y plane pair into a P/S Model 755 Nim Logic unit. A copy of each of the four individual plane pre-triggers is also sent to another set of P/S Model 755 to make a 3/4 or
4/4 plane coincidence (via a front-panel knob) which defines the production hodoscope pre-trigger. A copy of all the pre-triggers discussed above are sent to TDCs/Scalers via a NIM/ECL converter for timing and counting rate information. (See Figure \ref{fig:hms_hod_trg}) \\

\subsection{Calorimeter Pre-Trigger}
The HMS Calorimeter consists of four layers of lead blocks. Layers A and B read out a 26 PMT signals per layer (13 signals/side) while layers C and D
read out 13 signals/layer on one side. The first two layers form the Pre-Shower Counter while all four layers (A, B, C and D) form the Shower Counter.  
Each layer is read out in the Counting Room patch and fed into 50:50 splitters. One output of the splitter is fed to fADCs via a Ribbon-to-BNC converter (same as hodoscopes)
while the other output is sent to a P/S Model 740 NIM Linear FI/FO summing modules. Each side of a layer is summed first (hA+, hA-, hB+, hB-, hC and hD sums). The sums are fed into
a LeCroy Model 428F summing module where layers hA+/- and hB+/- are summed to form hA and hB sums. A copy of each layer sum is sent to fADCs. The Pre-Shower SUM is then made from the sum of layers A and B, while the Shower SUM is made by summing all four layers. A copy of the PreShower and Shower sums is also sent to fADCs. The Pre-Shower and Shower sums are then sent to a P/S Model 715
NIM Discriminator unit to form the PreShower Low/High (LO/HI) and Shower Low pre-triggers with thresholds \hPrShLo, \hPrShHi and \hSHLo, respectively with all gate widths set to 30 ns. A copy of the pre-triggers is sent to TDCs/Scaler modules for trigger timing and counting rate
information.
\newpage
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.5in, height=4.5in]{HMS_Cal_Trigger.pdf}
  \caption{HMS Calorimeter Electronics Diagram.}
  \label{fig:hms_cal_trg}
\end{figure}

\subsection{Gas \v{C}erenkov Pre-Trigger} \label{hms_cer_section}

\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=2.0in]{HMS_Cer_trig.pdf}
  \caption{HMS Gas \v{C}erenkov Electronics Diagram. Same electronics diagram applies for SHMS Gas \v{C}erenkov.}
  \label{fig:hms_gas_cer_trg}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=6.5in, height=2.1in]{HMS_Aero_trig.pdf}
  \caption{Aerogel \v{C}erenkovs Electronics Diagram. Same electronics diagram applies for SHMS Aerogel.}
  \label{fig:hms_aero_trg}
\end{figure}

\indent The HMS Aerogel \v{C}erenkov detector consists of a 120cm x 70cm rectangular aerogel tray coupled to a diffusion box\cite{hms_aero_article} and is located
between the second Drift Chamber and first set of hodoscope planes. The diffusion box has 8 PMTs on each side
which detect \v{C}erenkov light produced by interactions with the Aerogel material. The signals are sent directly to the Hall C Floor Patch Panel, and then
read out in the Counting Room patch and pass through a 50:50 splitter. One output is fed into an fADC module via a Ribbon-to-BNC converter. The other output is
sent to a summing module, and a copy of the sum is fed to an fADC. The sum is also sent to a NIM discriminator to form the Aerogel pre-trigger.
A copy of the discriminated signal is also sent to TDCs/Scalers via a NIM/ECL converter for trigger and counting rate information.

\subsection{HMS Single Arm Trigger}\label{ssec:hms_single_arm_sec}
\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=3.8in]{HMS_SingleArm_Trigger.pdf}
  \caption{HMS Single Arm Trigger Electronics Diagram}
  \label{fig:hms_one_arm_trg}
\end{figure}

@@

\section{SHMS Trigger Set-Up}
\indent The three planes (X1, Y1, X2) of scintillator arrays and the Quartz plane (Y2) will form part of the standard SHMS trigger configuration (See Figure \ref{fig:shms_stack}).  
Additional particle detectors may also be incorporated into the SHMS trigger as required by different experiments. The Noble Gas \v{C}erenkov \footnote{Noble Gas \v{C}er: e/$\pi$ separation $\geq$ 6 GeV/c} and
Calorimeter triggers will be used for e/$\pi$ separation, whereas the Heavy Gas\footnote{Heavy Gas \v{C}er: $\pi$/K separation above 3.4 GeV/c} and Aerogel \v{C}erenkov trigger will be used for $\pi$/K/p
separation\cite{SHMS_PID}.

\subsection{Hodoscope Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=4.3in]{SHMS_HODO_TRIGGER.pdf}
  \caption{SHMS Hodoscope Electronics Diagram.}
  \label{fig:shms_hod_trg}
\end{figure}

@@

\indent Once a pre-trigger has been made for each plane, they are sent to a NIM/ECL converter (Level Translator - Phillips Scientific (or P/S) Model 7126) via twisted pair cables to convert the ECL signal
(twisted pair) to a NIM signal. The NIM output is then sent to individual sets of a P/S Model 752 NIM Logic unit to adjust the widths of each of the plane pre-triggers as necessary before making a coincidence.
An X-Y hodoscope plane coincidence (S1 = S1X \textit{AND} S1Y, S2 = S2X \textit{AND} S2Y) is then made by feeding each hodoscope X-Y plane pair into a P/S Model 755 Nim Logic unit. A copy of each of the four
individual plane pre-triggers is also sent to another set of P/S Model 755 to make a 3/4 or 4/4 plane coincidence (via a front-panel knob) which defines the production hodoscope pre-trigger. A copy of all the
pre-triggers discussed above are sent to TDCs/Scalers via a NIM/ECL converter for timing and counting rate information. (See Figure \ref{fig:shms_hod_trg}) \\

>>>>

\usepackage{url}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{graphicx}
\usepackage[scientific-notation=true]{siunitx}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage[top=0.6 in, bottom=0.7 in, left=0.7in, right=0.7in]{geometry}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{vector}
\usepackage{amsmath}
\usepackage{empheq}
\usepackage{amssymb}
\usepackage{tgtermes}
\usepackage{gensymb}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{epstopdf}
\usepackage[singlelinecheck=false]{caption}
\usepackage[hyperfootnotes=false, pdfnewwindow, colorlinks=true]{hyperref}
\usepackage[title]{appendix}
\usepackage[section]{placeins}
\usepackage{float}

\AtBeginDocument{% optionally set colors to your liking
  \hypersetup{
    urlcolor=blue,
    citecolor=blue,
    linkcolor=blue,
  }%
}

@@

\newcommand{\shodthrs}{-30 mV }
\newcommand{\quartzthrs}{-60 mV }
\newcommand{\shodgate}{60 ns } 
\newcommand{\sngcthrs}{-50 mV}
\newcommand{\sngcgate}{30 ns}
\newcommand{\shgcthrs}{-50 mV}
\newcommand{\shgcgate}{30 ns}
\newcommand{\saerthrs}{-50 mV}
\newcommand{\saergate}{30 ns}
\newcommand{\comment}[1]{}

\title{\LARGE \bf{Hall C 12 GeV Trigger Set-Up}}

\author{Carlos Yero \\ email \href{mailto:cyero002@fiu.edu}{cyero002@fiu.edu}}

\maketitle

@@

\subsection{HMS Detector Hut}
The HMS Drift Chambers are read out through a VXS Crate (ROC3) in the detector hut electronics rack. The signals are carried through 16-channel ribbon cables
which are fed in various CAEN1190 TDC modules. The \textbf{TI} (or Trigger Interface) module at the front end of the crate distributes the readout
trigger throughout all modules in the crate and initiates data readout.
\begin{figure}[h!]
  \centering
  \includegraphics[width=5.0in, height=4.0in]{../HMS_stack.pdf}
  \caption{HMS detector stack.}
  \label{fig:hms_stack}
\end{figure}\\
\newpage
\noindent The rest of the HMS detector signals(Gas \c{C}erenkov, Hodoscope, Calorimeter) are
sent to the Hall C Floor Patch Panel via the hut Patch, with the exception
of the Aerogel, which is sent directly from the detector to the Floor Patch.
All the signals are then sent to the Counting Room Patch Panel to be processed
by the electronics. (See Figure \ref{fig:hms_patch})

\begin{figure}[h!]
  \centering
  \includegraphics[width=5.0in, height=4.0in]{../HMS_Hut_Rack.pdf}
  \caption{HMS detector hut electronic rack and patch panels.}
  \label{fig:hms_hut_rack}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=6.0in, height=2.0in]{../HMS_patch.pdf}
  \caption{HMS Patch Diagram from detectors to Counting Room.}
  \label{fig:hms_patch}
\end{figure}

\subsection{SHMS Detector Hut}
Similarly to the HMS Drift Chambers, the SHMS Drift Chambers are also read out by TDCs in a VXS Crate in the SHMS electronics hut (See Figure \ref{fig:shms_hut_rack2}).
The Shower Counter 224 signals are fed directly into the Flash ADC (fADC-250) modules in a VXS Crate (ROC4). The Pre-Shower signals (x14/side) pass
through a 50:50 splitter and a part is fed into fADCs. The other part is partially summed in the hut and sent via the hut patch panel to the
Counting Room Patch (See Figure \ref{fig:shms_hut_rack1}). The rest of the SHMS detector signals (HGC/NGC, Hodoscope, Aerogel) are sent to the Counting Room via the hut patch panel (See Figure \ref{fig:shms_patch}). 
\newpage
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.0in, height=4.0in]{../SHMS_stack.pdf}
  \caption{SHMS detector stack.}
  \label{fig:shms_stack}
\end{figure}
\begin{figure}[h!]
\begin{minipage}{.5\textwidth}
  \includegraphics[width=3.0in, height=4.0in]{../SHMS_Hut_Rack01.pdf}
  \captionof{figure}{SHMS Hut Patch Panel (left) and Electronics Rack (right).}
  \label{fig:shms_hut_rack1}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \includegraphics[width=2.0in, height=4.0in]{../SHMS_Hut_Rack02.pdf}
  \caption{SHMS Drift Chamber Electronics Rack.}
  \label{fig:shms_hut_rack2}
\end{minipage}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.8]{../SHMS_patch.pdf}
  \caption{SHMS Patch Diagram from detectors to Counting Room.}
  \label{fig:shms_patch}
\end{figure}

\subsection{Hall C Counting Room}
Once the detector signals arrive at the Counting Room Patch (See Figure \ref{fig:cr_racks_patch}A), they are processed by the NIM/CAMAC electronics (See Figure \ref{fig:cr_racks_patch}B) to form the single arm and coincidence triggers
for each spectrometer. The signals are also sent to ADCs/TDCs to determine energy and timing information for individual detectors as well as trigger TDC information.

\begin{figure}[h!]
  \includegraphics[scale=0.5]{../CR_patch_combined.png}
  \caption{(A) Counting Room Patch Panels (left 2 racks). (B) Counting Room main
    Electronic Racks for HMS/SHMS detectors (right 2 racks).}
  \label{fig:cr_racks_patch}
\end{figure}
\newpage
\section{HMS Trigger Set-Up} 
\indent The XY scintillator arrays (hodoscope planes) will form part of the standard HMS trigger configuration.  
Additional particle detectors may also be incorporated into the HMS trigger as required by different experiments. The Gas \v{C}erenkov and Calorimeter triggers will be
used for e/$\pi$ separation, whereas the Aerogel \v{C}erenkov trigger will be used for $\pi$/K/p separation.

\subsection{Hodoscope Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{hHODO_diagram.png}
  \caption{HMS Hodoscope Electronics Diagram. For original diagram, see Appendix \ref{appendix:AppxA1}}
  \label{fig:hHODO_diagram}
\end{figure}
\noindent Each hodoscope plane consists of an array of scintillator bars coupled to a PMT at each end (See Figure \ref{fig:hms_stack}), so each bar reads out two signals. As shown in Figure \ref{fig:hHODO_diagram},
for example, hodoscope plane h1X consists of 32 signals (16 bars) read out in the Counting House (CH) patch. Each side of the plane (x16 signals/side) is fed into a 64-channel input passive splitter (16 Ch./set).
One-third (33\%) of the signal amplitude is sent via a 16-channel ribbon cable to a 64 Ch. input Ribbon-to-BNC converter (16 Ch./set) which outputs are fed into a 16-channel NIM input flash ADC (fADC). The remaining two-thirds (66\%)
of the signal amplitude is sent to a 16-Ch. input CAMAC Discriminator unit. The HMS discriminators thresholds and gate widths were set to \hhodthrs and \hhodgate, respectively.\\
\indent The discriminated signals are sent via two ribbon-cable outputs to CAEN1190 TDCs/Scalers (daisy-chained) and to a LeCroy 4564 CAMAC Logic unit to form the plane pre-triggers.
The Logic Unit takes four sets of 16-Ch. input ribbon cable and forms a 16-fold OR for each set by default. Further boolean operations are done through the module backplane by connecting a twisted pair cable to the pin corresponding to the
desired boolean operation. For hodoscope plane pre-triggers, the boolean operations are as follows:
\begin{empheq}[box=\fbox]{align*}
& \text{h1X = h1X+ (16-fold OR) \textit{AND} h1X- (16-fold OR)} \\ 
& \text{h1Y = h1Y+ (10-fold OR) \textit{AND} h1Y- (10-fold OR)} \\
& \text{h2X = h2X+ (16-fold OR) \textit{AND} h2X- (16-fold OR)} \\ 
& \text{h2Y = h2Y+ (10-fold OR) \textit{AND} h2Y- (10-fold OR)} 
\end{empheq}

\indent Once a pre-trigger has been made for each plane, they are sent to a NIM/ECL converter (Level Translator - Phillips Scientific (or P/S) Model 7126) via twisted pair cables to convert the ECL signal (twisted pair)
to a NIM signal. The NIM output is then sent to individual sets of a P/S Model 752 NIM Logic unit to adjust the widths of each of the plane pre-triggers as necessary before making a coincidence. An X-Y hodoscope plane
coincidence (h1 = h1X \textit{AND} h1Y, h2 = h2X \textit{AND} h2Y) is then made by feeding each hodoscope X-Y plane pair into a P/S Model 755 Nim Logic unit. A copy of each of the four individual plane pre-triggers is also sent to another set of P/S Model 755 to make a 3/4 or
4/4 plane coincidence (via a front-panel knob) which defines the production hodoscope pre-trigger. A copy of all the pre-triggers discussed above are sent to TDCs/Scalers via a NIM/ECL converter for timing and counting rate information. (See Figure \ref{fig:hHODO_diagram}) \\

\subsection{Calorimeter Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{hCAL_diagram.png}
  \caption{HMS Calorimeter Electronics Diagram. For original diagram, see Appendix \ref{appendix:AppxA2}}
  \label{fig:hCAL_diagram}
\end{figure}
\noindent The HMS Calorimeter consists of four layers of lead blocks. Layers A and B read out a 26 PMT signals per layer (13 signals/side) while layers C and D
read out 13 signals/layer on one side. The first two layers form the Pre-Shower Counter while all four layers (A, B, C and D) form the Shower Counter.  
Each layer is read out in the Counting Room patch and fed into 50:50 splitters. One output of the splitter is fed to fADCs via a Ribbon-to-BNC converter (same as hodoscopes)
while the other output is sent to a P/S Model 740 NIM Linear FI/FO summing modules. Each side of a layer is summed first (hA+, hA-, hB+, hB-, hC and hD sums). The sums are fed into
a LeCroy Model 428F summing module where layers hA+/- and hB+/- are summed to form hA and hB sums. A copy of each layer sum is sent to fADCs. The Pre-Shower SUM is then made from the sum of layers A and B, while the Shower SUM is made by summing all four layers. A copy of the PreShower and Shower sums is also sent to fADCs. The Pre-Shower and Shower sums are then sent to a P/S Model 715
NIM Discriminator unit to form the PreShower Low/High (LO/HI) and Shower Low pre-triggers with thresholds \hPrShLo, \hPrShHi and \hSHLo, respectively with all gate widths set to 30 ns. A copy of the pre-triggers is sent to TDCs/Scaler modules for trigger timing and counting rate
information.

\subsection{Gas \v{C}erenkov Pre-Trigger} \label{hms_cer_section}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{hCER_diagram.png}
  \caption{HMS Gas \v{C}erenkov Electronics Diagram. Same electronics diagram applies for SHMS Gas \v{C}erenkov. For original diagram, see Appendix \ref{appendix:Appx3} }
  \label{fig:hCER_diagram}
\end{figure}

@@

\indent The HMS Aerogel \v{C}erenkov detector consists of a 120cm x 70cm rectangular aerogel tray coupled to a diffusion box\cite{hms_aero_article} and is located
between the second Drift Chamber and first set of hodoscope planes. The diffusion box has 8 PMTs on each side
which detect \v{C}erenkov light produced by interactions with the Aerogel material. The signals are sent directly to the Hall C Floor Patch Panel, and then
read out in the Counting Room patch and pass through a 50:50 splitter. One output is fed into an fADC module via a Ribbon-to-BNC converter. The other output is
sent to a summing module, and a copy of the sum is fed to an fADC. The sum is also sent to a NIM discriminator to form the Aerogel pre-trigger.
A copy of the discriminated signal is also sent to TDCs/Scalers via a NIM/ECL converter for trigger and counting rate information.
The electronics diagram is the same as in Figure \ref{fig:hCER_diagram}. For the original diagram, see Appendix \ref{appendix:Appx4}

\newpage
\subsection{HMS Single Arm Pre-Trigger}\label{ssec:hms_single_arm_sec}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{HMS_SingleArm_diagram.png}
  \caption{HMS Single Arm Pre-Trigger Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx5}}
  \label{fig:HMS_SingleArm_diagram}
\end{figure}

@@

\newpage 
\section{SHMS Trigger Set-Up}
\indent The three planes (X1, Y1, X2) of scintillator arrays and the Quartz plane (Y2) will form part of the standard SHMS trigger configuration (See Figure \ref{fig:shms_stack}).  
Additional particle detectors may also be incorporated into the SHMS trigger as required by different experiments. The Noble Gas \v{C}erenkov \footnote{Noble Gas \v{C}er: e/$\pi$ separation $\geq$ 6 GeV/c} and
Calorimeter triggers will be used for e/$\pi$ separation, whereas the Heavy Gas\footnote{Heavy Gas \v{C}er: $\pi$/K separation above 3.4 GeV/c} and Aerogel \v{C}erenkov trigger will be used for $\pi$/K/p
separation\cite{SHMS_PID}.

\subsection{Hodoscopes Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{pHODO_diagram.png}
  \caption{SHMS Hodoscopes Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx6}}
  \label{fig:pHODO_diagram}
\end{figure}

@@

\indent Once a pre-trigger has been made for each plane, they are sent to a NIM/ECL converter (Level Translator - Phillips Scientific (or P/S) Model 7126) via twisted pair cables to convert the ECL signal
(twisted pair) to a NIM signal. The NIM output is then sent to individual sets of a P/S Model 752 NIM Logic unit to adjust the widths of each of the plane pre-triggers as necessary before making a coincidence.
An X-Y hodoscope plane coincidence (S1 = S1X \textit{AND} S1Y, S2 = S2X \textit{AND} S2Y) is then made by feeding each hodoscope X-Y plane pair into a P/S Model 755 Nim Logic unit. A copy of each of the four
individual plane pre-triggers is also sent to another set of P/S Model 755 to make a 3/4 or 4/4 plane coincidence (via a front-panel knob) which defines the production hodoscope pre-trigger. A copy of all the
pre-triggers discussed above are sent to TDCs/Scalers via a NIM/ECL converter for timing and counting rate information. (See Figure \ref{fig:pHODO_diagram}) \\

\subsection{Pre-Shower / Shower Calorimeter Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{pCAL_diagram.png}
  \caption{SHMS PreShower / Shower Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx7}}
  \label{fig:pCAL_diagram}
\end{figure}

The SHMS Pre-Shower consists of two sets of fourteen (PMT-coupled) lead blocks oriented perpendicular to the Shower Counter blocks\cite{shms_preSh_talk}. The initial sum was done in the SHMS Electronics Hut. PMT signals
from groups of four blocks were summed to form:
\begin{empheq}[box=\fbox]{align*}
  & \text{preSh SUM [1-4]: [1-4]L + [1-4]R} \\ 
  & \text{preSh SUM [5-8]: [5-8]L + [5-8]R} \\ 
  & \text{preSh SUM [9-12]: [9-12]L + [9-12]R} \\ 
  & \text{preSh SUM [13-14]: [13-14]L + [13-14]R}
\end{empheq}
The Shower counter consists of 224 lead blocks, each coupled to a PMT at the end. Becasue of the high channel density of this detector, its signals are sent
directly to the ROC 4 fADCs in the SHMS detector hut.
\newpage
\subsection{Heavy/Noble Gas \v{C}erenkov Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{pCER_diagram.png}
  \caption{SHMS Gas \v{C}erenkovs Electronics Diagram.}
  \label{fig:pCER_diagram}
\end{figure}
\noindent The SHMS Heavy Gas \v{C}erenkov detector consists of a 1 m long, 1.6 m in diameter cylindrical tank located between the first set of hodoscope planes and the Aerogel (See Figure \ref{fig:shms_stack}).
The tank is filled with a gas and has four thin spherical mirrors that focus the \v{C}erenkov photons towards four 5-inch PMTs\cite{shms_hgc_talk}. 
\newline
\indent The SHMS Noble Gas \v{C}erenkov detector consists of a 2 m long active length of Argon/Neon gas tank located before the first drift chamber (See Figure \ref{fig:shms_stack}).
The tank is filled with a gas and has four overlapping mirrors, that focus the \v{C}erenkov photons towards four 5-inch PMTs\cite{shms_ngc_talk}. 
For the original electronics diagram, See Appendix \ref{appendix:Appx4} on the HMS Gas \v{C}erenkov.
\subsection{Aerogel \v{C}erenkov Pre-Trigger}
\indent The SHMS Aerogel \v{C}erenkov detector consists of a 110 x 100 x 24.5 cm$^{3}$ rectangular aerogel tray coupled to a diffusion box. The diffusion box has seven 5-inch PMTs on each side
which detect \v{C}erenkov light produced by interactions with the Aerogel material\cite{shms_aero_article}. The detector is located between Heavy Gas \v{C}erenkov and second set of hodoscope planes
(See Figure \ref{fig:shms_stack}). The electronics diagram is the same as in Figure \ref{fig:pCER_diagram}. For the original diagram, see Appendix \ref{appendix:Appx4}
\newpage
\subsection{SHMS Single Arm Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{SHMS_SingleArm_diagram.png}
  \caption{SHMS Single Arm Trigger Electronics Diagram. For the original diagram, see Appendix \ref{appendix:Appx8}.}
  \label{fig:SHMS_SingleArm_diagram}
\end{figure}

\indent The SHMS single arm trigger is be formed exactly as the HMS single arm trigger, with the exception of the detectors involved. Compare the electronics diagrams in Figures \ref{fig:SHMS_SingleArm_diagram} and
\ref{fig:HMS_SingleArm_diagram}, and read Section \ref{ssec:hms_single_arm_sec} for a detailed description of the electronic diagrams. 
\newpage
\section{Coincidence Trigger Set-Up}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{Coin_diagram.png}
  \caption{Coincidence Trigger Electronics Diagram. For the original diagram, see Appendix \ref{appendix:Appx9}}
  \label{fig:Coin_TRG}
\end{figure}
\indent In coincidence mode, the HMS and SHMS pre-triggers are sent to a NIM Logic module, where the first pre-trigger that arrives will open
a coincidence time window during which the second pre-trigger may or may not arrive in time. This will determine whether two events that occurred in
each spectrometer are correlated with the event that originated at the target. If the coincidence pre-trigger is formed, a copy will be sent to
scalers and TDCs, while the other copy is sent to a NIM/ECL converter where three copies (may be more) are sent via twisted pairs to every TDC module and the TI, which will be the \textbf{ONLY} Trigger Master in coincidence mode. Additional copies (not shown) may need to be sent to the HMS readout crate (ROC 01) as well, since the coincidence pre-trigger\footnote{A copy of the coincidence pre-trigger will also need to be sent to the
  HMS/SHMS TDC modules present in the detector huts crates via the patch panel.} is common to all crates. \\
\indent Once the coincidence pre-trigger is processed by the TI Master in ROC 02, it is sent to a NIM/ECL converter, where a copy is sent to scalers and TDCs, and the other copy is sent to the ``TRG'' NIM input in the TDC front panel, and daisy-chained with other TDCs in ROC02 and ROC01. The remaining crates in the HMS/SHMS huts receive a copy of the accepted trigger via fiber optics lines running from the TI Master to the TIs in all other crates. The TIs then distribute the accepted trigger to all modules of their respective crates. 
\newpage
\onecolumn
\bibliography{trigger_v2}
\bibliographystyle{acm}
